---
title: "Stat151A Lab 6"
output: html_document
date: "2024-02-21"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
```


# Distributions in R
## Probability density functions
```{r}
# Evaluate the probability density function of each distribution
dnorm(1, mean = 0, sd = 1)  # normal distribution  
dexp(1.5, rate = 1)  # exponential distribution
dt(2, df = 5)  # t-distribution
dunif(0.3, min = 0, max = 1)  # uniform distribution
dbinom(2, size = 5, prob = 0.3)  # binomial distribution
dpois(1, lambda = 2)  # poisson distribution
```

### Exercise 1
Plot the density function of the normal distribution whose mean is 2 and 
variance is 4.
```{r}
# Leave your answer below
x <- seq(-2, 6, by = 0.01)
# Calculate the density function
y <- dnorm(x_values, mean = 2, sd = 4)
plot(x, y)

# Plot the density function
cbind(x, y) %>% as.data.frame() %>%
  ggplot() + 
  geom_point(aes( x = x, y = y)) +
  theme_
```


## Cumulative distribution functions
Recall that the cumulative distribution function of a random variable $X$ is 
$F(x) = \mathbb{P}(X \le x)$.
```{r}
# Evaluate the cumulative density function of each distribution
pnorm(1, mean = 0, sd = 1)  # normal distribution  
pexp(1.5, rate = 1)  # exponential distribution
pt(2, df = 5)  # t-distribution
punif(0.3, min = 0, max = 1)  # uniform distribution
pbinom(2, size = 5, prob = 0.3)  # binomial distribution
pbinom(2.5, size = 5, prob = 0.3)  # binomial distribution
ppois(1, lambda = 2)  # poisson distribution
ppois(1.5, lambda = 2)  # poisson distribution
```

### Exercise 2
Compute the following probabilities for $X \sim \text{Exp}(2)$.
(1) $\mathbb{P}(X < 2)$
(2) $\mathbb{P}(X \le 2)$
(3) $\mathbb{P}(X > 1)$
(4) $\mathbb{P}(X \ge 1)$
(5) $\mathbb{P}(1 < X < 2)$
(6) $\mathbb{P}(1 \le X < 2)$
```{r}
# Leave your answer below
pexp(2, rate = 2)
pexp(2, rate = 2)
1 - pexp(1, rate = 2)
1 - pexp(1, rate = 2)
pexp(2, rate = 2) - pexp(1, rate = 2)
pexp(2, rate = 2) - pexp(1, rate = 2)
```

### Exercise 3
Compute the following probabilities for $X \sim \text{Binomial}(8, 0.4)$.
(1) $\mathbb{P}(X < 6)$
(2) $\mathbb{P}(X \le 6)$
(3) $\mathbb{P}(X > 3)$
(4) $\mathbb{P}(X \ge 3)$
(5) $\mathbb{P}(3 < X < 6)$
(6) $\mathbb{P}(3 \le X < 6)$
```{r}
# Leave your answer below
pbinom(5, 8, 0.4)
pbinom(6, 8, 0.4)
1 - pbinom(3, 8, 0.4)
1 - pbinom(2, 8, 0.4)
pbinom(5, 8, 0.4) - pbinom(3, 8, 0.4)
pbinom(5, 8, 0.4) - pbinom(2, 8, 0.4)
```


## Quantile functions
Recall that the quantile function $q_X$ of a random variable $X$ is defined as 
follows. For each $\alpha$, 
$$q_X(\alpha) = x$$
if 
$$P(X \le x) = \alpha.$$
```{r}
# Evaluate the quantile function of each distribution
qnorm(0.95, mean = 0, sd = 1)  # normal distribution  
qexp(0.95, rate = 1)  # exponential distribution
qt(0.025, df = 5)  # t-distribution
qunif(0.1, min = 0, max = 1)  # uniform distribution
# The smallest value x s.t. F(x) >= p where F is the cdf
qbinom(0.3, size = 5, prob = 0.3)  # binomial distribution
qbinom(0.5, size = 5, prob = 0.3)  # binomial distribution
qpois(0.3, lambda = 2)  # poisson distribution
qpois(0.4, lambda = 2)  # poisson distribution
```

### Exercise 4
Let $X$ be the exponential distribution whose rate parameter is 3 and $Y$ be the
t-distribution whose degree of freedom is 5. Find $u$ in each case below.
(1) $\mathbb{P}(X \le u) = 0.2$
(2) $\mathbb{P}(X \ge u) = 0.1$
(3) $\mathbb{P}(Y \le u) = 0.05$
(4) $\mathbb{P}(Y \ge u) = 0.3$
(5) $\mathbb{P}(-u \le Y \le u) = 0.9$
(6) $\mathbb{P}(0 \le Y \le u) = 0.3$ 
```{r}
# Leave your answer below
qexp(0.2, rate = 3)
qexp(1 - 0.1, rate = 3)
qt(0.05, df = 5)
qt(1 - 0.3, df = 5)
-qt((1 - 0.9) / 2, df = 5)
qt(pt(0, df = 5) + 0.3, df = 5) # qt(0.2, df = 5)
```


## Random sampling
```{r}
set.seed(2024)
# Draw random samples from each distribution
rnorm(3, mean = 0, sd = 1)  # normal distribution  
rexp(2, rate = 1)  # exponential distribution
rt(5, df = 5)  # t-distribution
rbinom(4, size = 5, prob = 0.3)  # binomial distribution
rpois(2, lambda = 2)  # poisson distribution
runif(6, min = 0, max = 1)  # uniform distribution
```

### Exercise 5
Draw 3000 samples from $\text{Binomial}(400, 0.1)$ and plot a histogram of 
samples in a density scale. Compare it with the probability density function of
$N(40, 6^2)$.
```{r}
# Leave your answer below
set.seed(2024)
samples <- rbinom(3000, 400, 0.1)
hist(samples, freq = FALSE, breaks = 40)
x <- seq(20, 60, by = 0.5)
y <- dnorm(x, mean = 40, sd = 6)
lines(x, y)
```

### Exercise 6
Let $X$ be the standard normal distribution. Estimate the value of 
$\mathbb{E}[X^4]$ by drawing samples from the distribution.
```{r}
# Leave your answer below
samples <- rnorm(100000, mean = 0, sd = 1)
mean(samples ** 4)
```
